Clustering Beat-Chroma Patterns\\ in a Large Music Database

### 2010-03-24 Paper/Paper.tex
dataset now always in one word

Note that none of this information (segments, beats, bars)
can be assumed perfectly accurate.
%(\eg in \cite{Barrington2009a} authors obtain a better
%song segmentation than the Echo Nest). 
In practice, we have found them reasonable, 
and given the size of the data set, any rare imperfections or noise
can be diluted to irrelevance by the good examples.  
We also believe that patch sizes based on a number of beats or bars are more
meaningful than an arbitrary time length. This is discussed further in 
Section \ref{ssec:segment}.

@@

For testing, we made use of low quality (32kbps MP3) audio 
from the uspop2002 data set \cite{uspop2002}.  The collections
includes a wide variety of popular music from 10 different sub-genres.
% \footnote{We also tried to use Tzanetakis dataset,
% %\cite{Tzanetakis2002a},
% but the $30$ seconds segments seemed hard to analyze by the Echo Nest API, 
% most did not contain any bar.}.
We use this data to measure how well a codebook learned on
the Cowbell data set can represent new songs.  
We obtained Echo Nest features for the $8651$ songs contained in the data set.

\subsection{Setting}\label{ssec:setting}
We train codebooks based on beat-chroma patches composed of 4 or 8
beats (one or two bars).
Codebooks over the Cowbell
data set are trained using the VQ algorithm (Algorithm \ref{algo:vq}). We used a
learning rate of $\ell=0.01$ for $200$ iterations over the data.
%
Given these codebooks, we quantize the test set by assigning each
patch to a single codeword.  We report the average
distance between patches and their encodings and describe the usage
frequency of the different codes. %, \ie the size of the cluster around a code.

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.8\columnwidth]{data_sizes}
\end{center}
\caption{\small{Distortion for a codebook of size $100$ encoding one bar
at a time, a bar being represented by $4$ chroma feature vector.
Therefore, a codeword or a pattern is of size $12$ x $4$ = $48$.
Distortion measure on the test data set. Data size tested range
from $0$ (just initialization) to $500$ thousand. Patterns were selected at
random from the dataset of approximately $3,7$ million patterns.
}}
\label{fig:data_sizes}
\end{figure}

@@

We apply our codebook to the artist recognition task.
Our data set, ``artist20'', encompass $1402$ songs from $20$ artists, 
mostly rock and pop of different subgenres. Results exist in the
literature for that set \cite{Ellis2007}. The best accuracy is
$59\%$, whereas using only chroma as a representation yields an
accuracy of $33\%$. The latter results is obtained through a
Gaussian mixture model.

>>>>

Note that none of this information (segments, beats, bars)
can be assumed perfectly accurate.
%(\eg in \cite{Barrington2009a} authors obtain a better
%song segmentation than the Echo Nest). 
In practice, we have found them reasonable, 
and given the size of the dataset, any rare imperfections or noise
can be diluted to irrelevance by the good examples.  
We also believe that patch sizes based on a number of beats or bars are more
meaningful than an arbitrary time length. This is discussed further in 
Section \ref{ssec:segment}.

@@

For testing, we made use of low quality (32kbps MP3) audio 
from the uspop2002 dataset \cite{uspop2002}.  The collections
includes a wide variety of popular music from 10 different sub-genres.
% \footnote{We also tried to use Tzanetakis dataset,
% %\cite{Tzanetakis2002a},
% but the $30$ seconds segments seemed hard to analyze by the Echo Nest API, 
% most did not contain any bar.}.
We use this data to measure how well a codebook learned on
the Cowbell dataset can represent new songs.  
We obtained Echo Nest features for the $8651$ songs contained in the dataset.

\subsection{Setting}\label{ssec:setting}
We train codebooks based on beat-chroma patches composed of 4 or 8
beats (one or two bars).
Codebooks over the Cowbell
dataset are trained using the VQ algorithm (Algorithm \ref{algo:vq}). We used a
learning rate of $\ell=0.01$ for $200$ iterations over the data.
%
Given these codebooks, we quantize the test set by assigning each
patch to a single codeword.  We report the average
distance between patches and their encodings and describe the usage
frequency of the different codes. %, \ie the size of the cluster around a code.

@@

\begin{figure}[htb]
\begin{center}
\includegraphics[width=.8\columnwidth]{data_sizes}
\end{center}
\caption{\small{Distortion for a codebook of size $100$ encoding one bar
at a time, a bar being represented by $4$ chroma feature vector.
Therefore, a codeword or a pattern is of size $12$ x $4$ = $48$.
Distortion measure on the test dataset. Data size tested range
from $0$ (just initialization) to $500$ thousand. Patterns were selected at
random from the dataset of approximately $3,7$ million patterns.
}}
\label{fig:data_sizes}
\end{figure}

@@

We apply our codebook to the artist recognition task.
Our dataset, ``artist20'', encompass $1402$ songs from $20$ artists, 
mostly rock and pop of different subgenres. Results exist in the
literature for that set \cite{Ellis2007}. The best accuracy is
$59\%$, whereas using only chroma as a representation yields an
accuracy of $33\%$. The latter results is obtained through a
Gaussian mixture model.
